\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

\title{HW3}
\author{Rachel Kantor, Caspar Popova, Melia Tomlinson}
\date{April 2022}

\begin{document}

\maketitle

\section{Problem i}

Theorem. $L$ is not regular.

\noindent Proof. We will show that $L$ is not regular by the Myhill-Nerode theorem by showing that $R_L$ has an infinite number of equivalence classes.

\noindent Let the set $S$ over $\Sigma *$ be $\{a^{4^k} | n \in \mathbf{N}\}$.

\noindent Consider any pair of strings in $S$, $s_x = a^{4^x}$ and $s_y = a^{4^y}$ where $x \neq y$.

\noindent Then we will show that, $\exists z \in \Sigma *$ such that $a^{4^x} z \in L$ and $a^{4^y} z \notin L$.

\noindent Let $z=a^{3 \cdot 4^x}$. Then:

\begin{align*}
    s_x z = a^{4^x} a^{3 \cdot 4^x} \\
    s_y z = a^{4^y} a^{3 \cdot 4^x} \\
    |s_x z| = 4^x + 3 \cdot 4^x = 4 \cdot 4^x  = 4^{x+1} \\
    |s_y z| = 4^y + 3 \cdot 4^x = 4^x ( 4^{y-x} - 3)
\end{align*}

\noindent We have shown that $|s_x z|$ is a power of $4$, therefore $s_x z \in L$. $|s_y x|$ is not a power of $4$ because of subtracting $3$. Therefore $s_y z \notin L$.

\noindent Since $s_x z \in L \wedge s_y z \notin L$, then for any pair of strings $s_x$ and $s_y$ over $\Sigma *$, $s_x$ and $s_y$ are distinguishable. Therefore, there are an infinite number of equivalence classes, therefore by the Myhill-Nerode theorem, $L$ is not regular.

\section{Problem ii}

Theorem. $L$ is regular.

\noindent Proof. We will show that $L$ is regular by the Myhill-Nerode theorem by showing that $R_L$ has a finite number of equivalence classes. 

\noindent Let there be $5$ equivalence classes over $\Sigma *$:

\begin{align*}
    \{a^{4k}  \ | \ k \in \mathbf{N} \}
    \\ \{a^{4k+1}  \ | \ k \in \mathbf{N} \}
    \\ \{a^{4k+2}  \ | \ k \in \mathbf{N} \}
    \\ \{a^{4k+3}  \ | \ k \in \mathbf{N} \}
    \\ \{ s \ | \ s \ \text{contains} \ b \}
\end{align*}

\noindent The equivalence classes are disjoint: the last equivalence is clearly disjoint from the first four because of $b$. The first four equivalence classes are disjoint because given a string $s$ s.t. $|s| = 4k+i$, then $|s| = 4k+j$ if and only if $i=j$.

\noindent Because the $5$ equivalence classes are disjoint, we have shown that $R_L$ has a finite number of equivalence classes, and therefore by the Myhill-Nerode theorem, $L$ is regular.

\end{document}

